calibration

### 2016-04-04 Rmd/calibration.Rmd
Updated calibration.Rmd to include analysis of SD/CV of observations by DOI and feature

p_root <- set_base_path("ecoscales")
p_dat <- fp(p_root, "external/data/calibration")

# Read in and clean-up calibration data
# cal <- data.table(read_excel(fp(p_dat, "merged_calibration.xlsx")))
cal <- data.table(read_excel(fp(p_dat, "merged_calibrationff.xlsx")))
cal[, DOI := gsub("DOI:|doi:|DOI: |doi: |DOI ", "", DOI)]
cal[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
dropnm <-  c("V30", "V31", "journal", "exclusion reason", "sensitivity")
cal[, c(dropnm) := NULL]

@@

# use NA in n_sites as indicator that study was excluded
cal[, inout := ifelse(is.na(n_sites), 0, 1)]  
# write.csv(cal, file = fp(p_dat, "merged_calibrationf.csv"))

# external fixes and recodes, bring back in
cal <- data.table(read_excel(fp(p_dat, "merged_calibrationf.xlsx")))
cal[, notes := NULL]
for(j in names(cal)[c(4:5, 8:24)]) set(cal, j = j, value = as.numeric(cal[[j]]))
# length(which(is.na(cal$feature)))

@@

```{r, warning = FALSE, fig.cap="Figure 1: Number of records extracted by each rater (observer) for each of the 20 papers in the calibration set."}
doimat <- t(doisum[, -1, with = FALSE])
cols <- RColorBrewer::brewer.pal(n = 6, "Paired")#[c(3:1, 4:6)]
cxs <- seq(3, 0.8, length.out = 6)
pchs <- c(rep(20, 5), 4)
par(mar = c(4, 4, 0.1, 0.1))
plot(c(1, 20), c(0, 10), pch = "", xlab = "Paper reviewed", 
     ylab = "N extractable records", mgp = c(1.45, 0.7, -0.2))
polygon(x = c(0, 21, 21, 0, 0), y = c(-1, -1, 11, 11, 0), col = "grey70")
# plot(doimat[1, ], pch = 20, xlab = "Paper reviewed", 
#      ylab = "N extractable records", cex = cxs[1], col = cols[1])
for(i in 1:6) points(doimat[i, ], col = cols[i], pch = pchs[i], cex = cxs[i])
legend(x = "topleft", pch = pchs, pt.cex = cxs, col = cols, 
       legend = paste0("Rater", 1:6), bty = "n")

@@

## Variance in extracted statistics
```{r}
varnms <- c("plot_res", "n_sites", "sampled_area", "t_btwn_samp", 
            "study_duration")
mus <- cal[inout == 1, lapply(.SD, mean, na.rm = TRUE), by = .(DOI, observer),             .SDcols = varnms]
vars <- mus[, lapply(.SD, sd, na.rm = TRUE), by = DOI, .SDcols = varnms]
musmu <- mus[, lapply(.SD, mean, na.rm = TRUE), by = DOI, .SDcols = varnms]

@@

@@

```{r}
par(mfcol = c(10, 4), mar = c(2, 2, 2, 2), mgp = c(1.5, 0.5, 0))

mus[, plot(plot_res), by = DOI]

Tim needs to fix paleo studies in calibration (and main set) giving too much credit to effective area.

Others need to fix their study duration figures. Paul on paleo one in calibration, Choi, Chang, etc in main set.

For N-sites, count sites only once if there are repeated measures taken at them, e.g. Jonathan in 10.1111/j.1466-8238.2009.00454.x for bird counts

>>>>

p_root <- set_base_path("ecoscales")
p_dat <- fp(p_root, "external/data/calibration")
```

```{r, eval=FALSE}
# Read in and clean-up calibration data
cal <- data.table(read_excel(fp(p_dat, "merged_calibration.xlsx")))
cal[, DOI := gsub("DOI:|doi:|DOI: |doi: |DOI ", "", DOI)]
cal[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
names(cal)
dropnm <-  c("V29")#, "V31", "journal", "exclusion reason", "sensitivity")
cal[, c(dropnm) := NULL]

@@

# use NA in n_sites as indicator that study was excluded
cal[, inout := ifelse(is.na(n_sites), 0, 1)]  
# write.csv(cal, file = fp(p_dat, "merged_calibrationf.csv"))
```

```{r}
# external fixes and recodes, bring back in
# cal <- data.table(read_excel(fp(p_dat, "merged_calibrationf.xlsx")))
cal <- data.table(read_excel(fp(p_dat, "merged_calibrationff.xlsx")))
names(cal)
cal[, c("notes", "V29") := NULL]
for(j in names(cal)[c(4:5, 10:24)]) {
  set(cal, j = j, value = as.numeric(cal[[j]]))
}
# length(which(is.na(cal$feature)))

@@

```{r, warning = FALSE, fig.cap="Figure 1: Number of records extracted by each rater (observer) for each of the 20 papers in the calibration set."}
doimat <- t(doisum[, -1, with = FALSE])
cols <- RColorBrewer::brewer.pal(n = 6, "Paired")#[c(3:1, 4:6)]
cxs <- seq(3, 0.8, length.out = 6)
pchs <- c(rep(20, 5), 4)
par(mar = c(4, 4, 0.1, 0.1))
plot(c(1, 20), c(0, 10), pch = "", xlab = "Paper reviewed", 
     ylab = "N extractable records", mgp = c(1.45, 0.5, 0), tcl = -0.2)
polygon(x = c(0, 21, 21, 0, 0), y = c(-1, -1, 11, 11, 0), col = "grey70")
# plot(doimat[1, ], pch = 20, xlab = "Paper reviewed", 
#      ylab = "N extractable records", cex = cxs[1], col = cols[1])
for(i in 1:6) points(doimat[i, ], col = cols[i], pch = pchs[i], cex = cxs[i])
legend(x = "topleft", pch = pchs, pt.cex = cxs, col = cols, 
       legend = paste0("Rater", 1:6), bty = "n")

@@

## Variance in extracted statistics
### By DOI
```{r}
varnms <- c("plot_res", "n_sites", "sampled_area", "t_btwn_samp", 
            "study_duration")
mus <- cal[inout == 1, lapply(.SD, mean, na.rm = TRUE), by = .(DOI, observer),             .SDcols = varnms]  # mean variable val per paper per observer
# variance and mean per paper across observers
vars <- mus[, lapply(.SD, sd, na.rm = TRUE), by = DOI, .SDcols = varnms] 
musmu <- mus[, lapply(.SD, mean, na.rm = TRUE), by = DOI, .SDcols = varnms]

@@

# calr[DOI == "10.1007/s00442-005-0338-3" & feature == 2]
```

### By feature (individual records) across DOIs
```{r, results='hold'}
CV <- function(x) sd(x) / mean(x)
rnms <- c("DOI", "feature", "observer", "plot_res", "n_sites", "t_btwn_samp",
          "study_duration")
calr <- cal[, rnms, with = FALSE]
# calr[feature != 99, ]
# calr[feature != 99, lapply(.SD, mean), by = .(DOI, feature), 
#      .SDcols = rnms[4:7]]
# calr[feature != 99, .N, by = .(DOI, feature)]
calr[feature != 99, lapply(.SD, CV), by = .(DOI, feature), 
     .SDcols = rnms[4:7]][, lapply(.SD, mean, na.rm = TRUE), 
                          .SDcols = rnms[4:7]]

# par(mfcol = c(10, 4), mar = c(2, 2, 2, 2), mgp = c(1.5, 0.5, 0))
# mus[, plot(plot_res), by = DOI]

@@

Tim needs to fix paleo studies in calibration (and main set) giving too much credit to effective area. [DONE]

Others need to fix their study duration figures. Paul on paleo one in calibration, Choi, Chang, etc in main set.  [Ahmed remaining]

For N-sites, count sites only once if there are repeated measures taken at them, e.g. Jonathan in 10.1111/j.1466-8238.2009.00454.x for bird counts [Chang, Ahmed remaining]
